---
title: "ResBaz Geospatial R Workshop (Empty)"
output: html_notebook
---

Run me before the workshop and ensure no errors are produced!
```{r Workshop setup}
#Run me once (must undo # but replace after execution)
#install.packages(c("dplyr", "terra", "sf", "rasterVis", "ggplot2", "ggspatial"))
```


```{r libraries}
#Convenience package
library(dplyr) #Piping and data manipulations

#Geospatial manipulations
library(sf) #Non-gridded data (point, polygon, line) manipulation
library(terra) #Gridded (primarily) data manipulation
 
#Mapping
library(ggplot2) #sf plotting
library(ggspatial) #mapping icons
library(rasterVis) #raster plotting package
```

Note - many of the simple operations you can perform with sf can also be done with terra (see terra::vect() objects). However, sf is purpose built for simple features (points, lines and polygons) and is integrated with terra so the two can be used together!

##sf package
###Exercise One
```{r Exercise One Loading sf data}
#Shapefile loaded with a relative path (relative to the working directory which is where ever the geospatialResBaz.proj file is)

#1. Read in nz_coastlines shapefile data 

#2. Plot these data


```

###Exercise Two
```{r Exercise Two - sf point data}
#1. Read in drosera.shp point data

#2. Transform these to NZGD2000 (ESPG 2193)

#3. Plot points (colour by species) and outline of NZ
```

###Exercise Three
```{r Exercise Three - ggspatial}
#1. Add to your map:
#Scale bar
#North arrow

#Plotting sf data

#Adding stuff to make the map pretty
 
#North arrow

#Saving the map ggsave?
#note: ggsave smooths jagged lines

```

###Exercise Four
```{r Exercise Four - Cropping}
#Defining a boundary box (a.k.a. extent)
stew.bbox <- st_bbox(c(xmin = 1155000, #Boundaries
                       ymin = 4723000,  
                       xmax = 1285000, 
                       ymax = 4820000), 
                     class = "bbox", #What it is
                     crs = 2193) #The ESPG code for the CRS

#Check out this site for a great tool to create bounding boxes by hand!http://bboxfinder.com/

#1. Crop NZ coastline and drosera data layer to match

#Hint - crop these separately!
#Cropping the points
#Cropping the coastline


#2. Plot the cropped sf objects
#plotting the cropped data

```

###Exercise Five
```{r Exercise Five - Leaflet}

#1. Read in the drosera.csv file (read.csv)

#2.  Convert df to a leaflet object

#3. Make the species as a pop-up value

#4. Open the viewer


```

##terra package
###Exercise Six
```{r Exercise Six - Loading Raster data}
#Raster loaded with a relative path 
#1. Load the “precip_warmQtr.tif”  file assign it to the name precip

#2. Load the “temp_meanColdQtr.tif” and assign it to temp_cold 

#3.  Plot these raster data - gg won't well work here!

```

###Exercise Seven
```{r Exercise Seven - Stacks and Projections}
#1. Combine the precip and temp_cold objects into a single object named rastStack

#2. Load “elevation49.tif” (not elevation.tif!) and assign the name elev49


#3. Try to add the elev49 object to rastStack

#Is there a problem?


```

###Exercise Eight
```{r Exercise Eight - Raster Algebra}
#1. Load “temp_meanWarmQtr” and assign it to temp_warm 

#2. Create a raster of the difference between temp_cold & temp_warm and assign it to temp_range


#3. Combine these into a single stack called temps

```

###Exercise Nine
```{r Exercise Nine - rasterVis plotting}
#1. Use levelplot() to plot temp_range

#2. Add contours to the map

#3. Use gplot to plot a single layer


#Extra code for multiplots and saving levelplot graphs!
library(patchwork)

p1 <- gplot(rastStack[[1]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "magma", 
                       na.value = "transparent",
                       name = names(rastStack[[1]])) +
  coord_equal()

p2 <- gplot(rastStack[[2]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "viridis", 
                       na.value = "transparent",
                       name = names(rastStack[[2]])) +
  coord_equal() 


p3 <- gplot(rastStack[[3]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "cividis", 
                       na.value = "transparent",
                       name = names(rastStack[[1]])) +
  coord_equal() 

#Plotting with patchwork
p1 + p2 + p3

#Saving raster plots created by levelplot
png("./graphs/precip_plot.png", width = 725, height = 1024)
levelplot(precip) 
dev.off()   


```

##Terra and sf combine!
```{r - Example of analysis, eval = FALSE }
clim.bg <- spatSample(temps, 10000) %>% 
  mutate(species = "Background")

#Pulling the sf data
drosera.df2 <- drosera.sf %>% #Starting with my sf object I loaded from a shapefile
  st_coordinates() %>% #Extracting the coordinates I want
  as.data.frame() %>% #Converting to a dataframe
  rename("longitude" = X , "latitude" = Y) #Renaming the columns for clarity

#Prepping the data 
clim.all <- terra::extract(temps, 
                           drosera.df2)[,-1] %>% #Pulling the values for each point from the rasters
  bind_cols(species = as.character(drosera.sf$species)) %>% #Adding on the species data I had so I know which obs is for which species
  bind_rows(clim.bg) #Adding on my background data


ggplot(clim.all,
       aes(y = temp_meanWarmQtr,
           x = temp_meanColdQtr,
           colour = species)) + #Setting up the plot
  
  geom_point(data = filter(clim.all, 
                           species == "Background"),
             alpha = 0.4) + #Plotting the background points
  
  geom_point(data = filter(clim.all, 
                           species != "Background"),
             alpha = 0.6) + #Plotting my species points
  
  scale_colour_manual(values = c("grey", "#802028", "#586880"), name = "Species") + #Colouring!
  
  theme_minimal() #Because I like it.
```
