---
title: "ResBaz Geospatial R Workshop"
output:
  html_document:
    df_print: paged
---

Run me before the workshop and ensure no errors are produced!
```{r Workshop setup}
#Run me once (must undo # but replace after execution)
#install.packages(c("dplyr", "terra", "sf", "rasterVis", "ggplot2", "ggspatial"))
```


```{r libraries}
#Convenience package
library(dplyr) #Piping and data manipulations

#Geospatial manipulations
library(sf) #Non-gridded data (point, polygon, line) manipulation
library(terra) #Gridded (primarily) data manipulation
 
#Mapping
library(ggplot2) #sf plotting
library(ggspatial) #mapping icons
library(rasterVis) #Raster plotting package
library(leaflet)
```
Note - many of the simple operations you can perform with sf can also be done with terra (see terra::vect() objects). However, sf is purpose built for simple features (points, lines and polygons) and is integrated with terra so the two can be used together!

##sf package
###Exercise One
```{r Exercise One Loading sf data}
#Raster loaded with a relative path (relative to the working directory which is where ever the geospatialResBaz.proj file is)

#1. Read in nz_coastlines shapefile data 
nz.sf <-  st_read("../data/polygons/nz_coastlines/nz_coastlines.shp")

#2. Plot these data
#Plotting
ggplot() +
  geom_sf(data = nz.sf) #sf plots sf objects and "data =.." is necessary!

#OR
plot(nz.sf)
```

###Exercise Two
```{r Exercise Two - sf point data}
#1. Read in nz_coastlines shapefile data 
#Reading in the data
drosera.sf <-  st_read("../data/point_data/drosera/drosera.shp") %>% 
  st_transform(2193)


#2. Plot these data
#Plotting the data together
ggplot() +
  geom_sf(data = nz.sf) +
  geom_sf(data = drosera.sf, 
          aes(colour = species),
          alpha = 0.5)
```

###Exercise Three
```{r Exercise Three - ggspatial}
#1. Add to your map:
#Scale bar
#North arrow

#Plotting sf data
ggplot() +
  geom_sf(data = nz.sf) +
  geom_sf(data = drosera.sf, 
          aes(colour = species), alpha = 0.5) +
  #Adding stuff to make the map pretty
  annotation_scale(location = "br") + #Scale
  annotation_north_arrow(location = "tl", 
                         which_north = "true", 
                         style = north_arrow_fancy_orienteering) + #North arrow
  theme_minimal() #Looks nice..

#Saving the map
ggsave("../graphs/exampleMap.png")
#note: ggsave smooths jagged lines
```

###Exercise Four
```{r Exercise Four - Cropping}

#Defining a boundary box (a.k.a. extent)
stew.bbox <- st_bbox(c(xmin = 1155000, #Boundaries
                       ymin = 4723000,  
                       xmax = 1285000, 
                       ymax = 4820000), 
                     class = "bbox", #What it is
                     crs = 2193) #The ESPG code for the CRS

#1. Crop NZ coastline and drosera data layer to match
#Cropping the points
dros_stewart.sf <- st_crop(drosera.sf, stew.bbox)

#Cropping the coastline
coast_stewart.sf <- st_crop(nz.sf, stew.bbox)

#2. Plot the cropped sf objects
#plotting the cropped data
ggplot() + 
  geom_sf(data = coast_stewart.sf) + 
  geom_sf(data = dros_stewart.sf, 
          aes(colour = species))

```

###Exercise Five
```{r Exercise Five - Leaflet}

#1. Read in the drosera.csv file (read.csv)
#Reading in csv data
drosera.df <- read.csv("../data/point_data/drosera.csv")

#2.  Convert df to a leaflet object
#To show
head(drosera.df)

#3. Make the species as a pop-up value
#Using leaflet
drosera.leaf <- leaflet(drosera.df) %>%
  addTiles() %>% 
  addCircleMarkers(lng = ~long, 
                   lat = ~lat, 
                   popup = ~species)

#4. Open the viewer
drosera.leaf

```

##terra package

###Exercise Six
```{r Exercise Six - Loading Raster data}
#1. Load the “precip_warmQtr.tif”  file assign it to the name precip
#Raster loaded with a relative path 
precip <- rast("../data/rasters/precip_warmQtr.tif") 

#2. Load the “temp_meanColdQtr.tif” and assign it to temp_cold 
temp_cold <- rast("../data/rasters/temp_meanColdQtr.tif")

#3.  Plot these raster data
#Plotting
plot(precip)
plot(temp_cold)

```

###Exercise Seven
```{r Exercise Seven - Stacks and Projections}
#1. Combine the precip and temp_cold objects into a single object named rastStack
#Combining two rasters
rastStack <- c(precip, temp_cold)

# rastStack <-  rast(c("../data/rasters/precip_warmQtr.tif", 
#                        "../data/rasters/temp_meanColdQtr.tif"))

rastStack[[1]]
rastStack[["precip_warmQtr"]]

#2. Load “elevation49.tif” (not elevation.tif!) and assign the name elev49
#Reading in elev49
# elev49 <- rast("../data/rasters/elevation_49.tif")
# 
# #3. Try to add the elev49 object to rastStack
# #Trying to add
# rastStack <- c(rastStack, elev49)

#Reprojecting

#This process can take a few minutes - more for large files!

# elev <- project(elev, #Problem raster
#                 rastStack, #Target raster
#                 filename = "../data/rasters/elev.tif", #This will write it as a new tif file
#                 overwrite = TRUE) #Allows files with the same name to be over-written

#If processing is too slow.
elev <- rast("../data/rasters/elevation.tif")

#Adding them all together in a single stack
rastStack <- c(rastStack, elev)


```

###Exercise Eight
```{r Exercise Eight - Raster Algebra}
#1. Load “temp_meanWarmQtr” and assign it to temp_warm 
#Reading in the new layer
temp_warm <- rast("../data/rasters/temp_meanWarmQtr.tif")

#2. Create a raster of the difference between temp_cold & temp_warm and assign it to temp_range
temp_range <-  temp_warm - temp_cold
plot(temp_range)

#3. Combine these into a single stack called temps
#Combining data
temps <- c(temp_warm, temp_cold, temp_range)

#Extra stuff as an example
#This create a raster of TRUE/FALSE (coded as 0 and 1) showing the cells that have a value as less than 10 as true
temp_range.lt10 <- temp_range < 10
plot(temp_range.lt10)

```

###Exercise Nine 
```{r Exercise Nine - rasterVis plotting}
#1. Use levelplot() to plot temp_range
levelplot(temp_range)

#2. Add contours to the map
levelplot(temp_range, contour = TRUE)

#3. Use gplot to plot a single layer
#rasterVis plotting 1
gplot(temp_range) +
  geom_raster(aes(fill = value)) +
  facet_wrap(~variable) +
  scale_fill_viridis_c(option = "magma", 
                      na.value = "transparent") +
  coord_equal()

#Extra goodies!!!

#patchwork
library(patchwork)
#Multiple plots with varying axes
p1 <- gplot(rastStack[[1]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "magma",
                       na.value = "transparent",
                       name = names(rastStack[[1]])) +
  coord_equal()

p2 <- gplot(rastStack[[2]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "viridis",
                       na.value = "transparent",
                       name = names(rastStack[[2]])) +
  coord_equal()


p3 <- gplot(rastStack[[3]]) +
  geom_raster(aes(fill = value)) +
  scale_fill_viridis_c(option = "cividis",
                       na.value = "transparent",
                       name = names(rastStack[[1]])) +
  coord_equal()

#Plotting with patchwork
p1 + p2 + p3

#Plotting logical vectors
gplot(temp_range.lt10) +
  geom_raster(aes(fill = as.logical(value))) +
  scale_fill_manual(values = c('red', 'blue'),
                    na.value = "transparent",
                    name = "Diff < 10 °C") +
  labs(x = 'Longitude', y = 'Latitude') +
  theme_minimal() +
  coord_equal()

```

##Terra and sf combine!
```{r - Example of analysis }
#Getting a background sample
clim.bg <- spatSample(temps, 10000) %>% 
  mutate(species = "Background")

#Pulling the sf data
drosera.df2 <- drosera.sf %>% #Starting with my sf object I loaded from a shapefile
  st_coordinates() %>% #Extracting the coordinates I want
  as.data.frame() %>% #Converting to a dataframe
  rename("longitude" = X , "latitude" = Y) #Renaming the columns for clarity

#Prepping the data 
clim.all <- terra::extract(temps, 
                           drosera.df2)[,-1] %>% #Pulling the values for each point from the rasters
  bind_cols(species = as.character(drosera.sf$species)) %>% #Adding on the species data I had so I know which obs is for which species
  bind_rows(clim.bg) #Adding on my background data


ggplot(clim.all,
       aes(y = temp_meanWarmQtr,
           x = temp_meanColdQtr,
           colour = species)) + #Setting up the plot
  
  geom_point(data = filter(clim.all, 
                           species == "Background"),
             alpha = 0.4) + #Plotting the background points
  
  geom_point(data = filter(clim.all, 
                           species != "Background"),
             alpha = 0.6) + #Plotting my species points
  
  scale_colour_manual(values = c("grey", "#802028", "#586880"), name = "Species") + #Colouring!
  
  theme_minimal() #Because I like it.

#Saving
#ggsave("../graphs/exampleDroseraPlot.png", width = 7.6, height = 5.2)

```

##Extra code for raster manipulation
```{r - Extra code, eval = FALSE}
#RASTER MANIPULATIONS
#Reading in a spat vector (terra polygon)
veg <- vect("./data/polygons/native_veg/native_veg.shp")

plot(veg)

#Masking
maskStack <- mask(temp_warm, veg)
plot(maskStack[[1]])

#Cropping
#Defining the extent of the cropping frame
cropBox <- c(1500000, 1650000, 
             5100000, 5200000) %>% 
  ext() #Making it a ext terra object

#Cropping it
tempCrop <- crop(temp_warm, cropBox)
plot(tempCrop)

# Workshop material prep 

#Pulling data from GBIF
xx <- rgbif::occ_search(scientificName = c("Drosera binata",
                                           "Drosera spatulata"),
                        country = 'NZ',
                        hasCoordinate = TRUE)

#Point data
drosera.df <-bind_rows(xx$`Drosera binata`$data,
                       xx$`Drosera spatulata`$data) %>% 
  dplyr::select(c(species, 
                  decimalLatitude, 
                  decimalLongitude)) %>% 
  filter(!is.na(decimalLatitude))

#Converting to sf object
drosera.sf <- drosera.df %>% 
  st_as_sf(coords = c("decimalLongitude", 
                      "decimalLatitude")) %>% 
  st_set_crs(4326) %>% #WSG84
  st_transform(crs = 2193) %>% #NZGD2000
  st_join(nz.sf, join = st_within) %>% #Figuring which points are erroneous/NA
  filter(!is.na(TARGET_FID)) %>%  #Removing these from the dataset
  st_transform(4326)

st_write(drosera.sf, "./data/point_data/drosera.shp")

#Converting back to a dataframe
drosera.df <- drosera.sf %>% 
  st_transform(4326) %>% 
  st_coordinates() %>%
  as.data.frame() %>% 
  bind_cols(drosera.sf$species) %>% 
  rename(lat = Y, long = X, species = ..3)

#Writing to file
write.csv(drosera.df, "./data/point_data/drosera.csv")
```



